---
title: "OEC R Package - First Steps"
author: "Mauricio Vargas S."
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{OEC R Package - Complete User Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction

`oec` is an [Observatory of Economic Complexity](http://atlas.media.mit.edu/en/) API wrapper and utility program. Since v1.0 its available on [CRAN](https://cran.r-project.org/package=oec).

## Data sources

All of the product data shown on the site is classified using either SITC (Standard International Trade Classification) or HS (Harmonized System).

For historical SITC classification data (1962 - 2000), we use data from [The Center for International Data from Robert Feenstra](http://cid.econ.ucdavis.edu/). For more recent data (2001 - 2016), we use data provided by [UN Comtrade](http://comtrade.un.org/).

## How to use the package?

The easiest way to use this package, and R in general, is to use [RStudio IDE](https://www.rstudio.com/) that allows users to create projects that do not depend on absolute paths.

The package is fully documented [here](examples/oec.pdf). In this vignette we provide some examples that do not replace the documentation.

### Basic use

The first step is to install the package and explore the country codes to start downloading data.

```{r functions1, eval = FALSE}
# Install the package from CRAN (recommended)
install.packages("oec")

# Or install the package from Github (this option or CRAN option, not both)
#install.packages("devtools") ## if needed
library(devtools)
install_github("observatory-economic-complexity/oec-r")

# Load the package
library(oec)

# Explore the countries list
countries_list
```

Now we are ready to download data and provide an answer to questions like *What does Andorra export?*

```{r functions2, eval = FALSE}
# What does Andorra export? (2015, HS92 4 characters)
getdata("and", "all", 2015)

# What does Andorra export? (2015, SITC rev2 4 characters)
getdata("and", "all", 2015, 2)

# What does Andorra export? (2014 and 2015, HS92 4 characters)
getdata_interval("and", "all", 2014, 2015)

# What does Andorra export? (2014 and 2015, SITC rev2 4 characters)
getdata_interval("and", "all", 2013, 2015, 2)
```
